---
title: STAT 5443, HW1
author: Ariel Mundo
date: 9/15/2020
output:
    #df_print: paged
  prettydoc::html_pretty:
    theme: cayman
    #highlight: github
---


```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(prettydoc)
```
***
 Disclaimer: This is the first time I try using RMarkdown, please excuse any typing errors.


#### PROBLEM 1

```{r, PROBLEM 1}
x<-c(3, 4, 12, 19, 23, 48, 50, 61, 63, 78)
primes<-c(2, 3, 5, 7, 11, 13, 17, 19, 23, 29, 31, 37, 41, 
          43, 47, 53, 59, 61, 67, 71, 73, 79, 83, 89, 97)     
j<-rep(0,length(x)) #variable to keep track of index of non prime numbers in x
for (i in 1:length(x)){
  p<-x[i]==primes #logical comparison of each element in x to prime numbers
  if(all(p==FALSE)){
                    j[i]<-i} #if all the elements on p are false, the number in x[i]
                              #is a non-prime integer, keep track of it using j
}
print(x[j]) #print the non-prime numbers


```
***
#### PROBLEM 2.a

According to the definition $f(n)=O(g(n))$ if the limit equals a constant $k>0$. Taking the limit of:
$\lim_{x \to \infty}\frac {x^3} {x^4}$ using L'Hopital rule renders: \[ \lim_{x \to \infty}\frac{3}{4x} \neq 0\], and therefore $x^{3}$ is not $\Theta(x)^{4}$

PROBLEM 2.b

Rewriting the expression: \[\lim_{n \to \infty}(\frac{\log n} {n^{1/k}})^k\], and using L'Hopital rule: \[\lim_{n \to \infty}\frac {k}{n^{1/k}}=0\] and therefore $(\log(n))^{k}=O(n)$

***
PROBLEM 3.a
```{r, PROBLEM 3.a}
###selection sort####
selection_sort<-function(A){
  r<-length(A)
  for (i in 1:(r-1)){
    index<-i #tracks smallest number in the original array, updated accordingly 
    for (j in (i+1):r){
      if (A[j]<A[index]){
        index<-j #updated if something smaller is found
      }
    }
  if (index!=i){ #if updated, swap the values
    temp<-A[i] #move the number
    A[i]<-A[index] #put in the first position
    A[index]<-temp #put the following number in the next position
      }
    }
  return(A)
}
```

PROBLEM 3.b ELAPSED TIMES FOR EACH ALGORITHM
```{r, PROBLEM 3.b, echo=FALSE, results='hide'}
#####merge sort####
MergeSort <- function(x){
  if(length(x) == 1) {
    res <- x
  }else{
    # split x in 2 and call mergesort on both parts
    i <- floor(length(x)/2)
    s1 <- c(MergeSort(x[1:i]), Inf)
    s2 <- c(MergeSort(x[-(1:i)]), Inf)
    res <- rep(NA_real_, length(x))
    # merge the responses in ascending order
    i1 <- i2 <- 1
    for (i3 in 1:length(res)){
      if (s1[i1] <= s2[i2]) {
        res[i3] <- s1[i1]
        i1 <- i1 + 1
      } else {
        res[i3] <- s2[i2]
        i2 <- i2 + 1
      }
    }      
  } 
  return(res)
}
p<-rnorm(1000)
ptm<-proc.time()
selection_sort(p)
t1<-proc.time()-ptm
ptm<-proc.time()
MergeSort(p)
t2<-proc.time()-ptm
```

```{r}
print(t1[['elapsed']]) ## selection sort
print(t2[['elapsed']]) ## merge sort
```



```{r, PROBLEM 3.c}
reps <- 100
time1 <- time2 <- rep (0,reps )
for (i in 1: reps ){
  # obtain the cpu time from the i - th random sample
  # and save the results as , say , t1 and t2
  y = rnorm (1000) # generated sample of size 1000
  ptm<- proc.time()
  MergeSort (y) # merge sort algorithm
  t11 <- proc.time()-ptm
  time1[i]<- t11[['elapsed']]
  ptm <- proc.time()
  selection_sort(y) # Selection sort algorithm
  t22 <-proc.time()-ptm
  time2[i]<- t22[['elapsed']]
}

mean1<-mean(time1)
var1<-var(time1)
print(paste('Merge sort mean time:',mean1))
print(paste('Merge sort variance:',var1))
mean2<-mean(time2)
var2<-var(time2)
print(paste('Selection sort mean time:',mean2))
print(paste('Selection sort variance:',var2))
ratio<-mean2/mean1
print(paste('Ratio of selection sort time to merge sort time (mean)',ratio))
```
#### PLOTTING THE RESULTS

```{r}
plot(time1)
plot(time2)
```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
#### *The values do agree, for a single call to each function there is no significant difference, but upon recursive calling, mergesort is about 4 times faster than selection sort.*
</div>

***
#### BONUS PROBLEM

BONUS PROBLEM a

From the problem statement, this corresponds to:

$5T(n/2)+O(n^{1})$ where $a=5, b=2, d=1$ therefore $d<\log_{2}(5)$ and according to the Master Theorem:
\[T(n)=O(n^{\log_{2}5})\]

BONUS PROBLEM b

From the problem statement: $T(n)=2T(n-1)+k$, but this form has no correspondence in the Master Theorem. However, this corresponds to the "Towers of Hanoi" problem. In this problem, for 1 disk, $T(n)=1$. Using $k=1$ for two,three, four and five disks:

\[T(2)=2*1+1=3\]
\[T(3)=2*3+1=7\]
\[T(4)=2*7+1=15\]
\[T(5)=2*15+1=31\]

By intuition, this corresponds to $2^{n}-1$. Since in big $O$ notation the constants are dropped, the time complexity for this algorithm would be: \[T(n)=O(2^{n})\]


BONUS PROBLEM c

From the problem statement: $T(n)=9T(n/3)+O(n^{2})$ where $a=9, b=3, d=2$. From the Master Theorem: \[2=\log_{3}9\] 
Therefore: \[T(n)=O(n^{2} \log n)\]