---
title: STAT 5443, HW2
author: Ariel Mundo
date: 9/29/2020
output:
    prettydoc::html_pretty:
    theme: cayman
    #css: styles.css
---
<style type="text/css">

p {
   font-size: 10px;
}
code.r{
  font-size: 10px;
}
pre {
  font-size: 10px
}
</style>



```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(prettydoc)
```

#### PROBLEM 1

##### PART A

Since $p(x)=\sin(x)$ the $CDF$ is calculated as:

\[\int_0^{x} \sin w~dw\] which yields:
\[F(x)=1-\cos(x)\] 
and using the inverse CDF Transform:

\[u=1-\cos(F^{-1}(u))\] 
\[F^{-1}(u)=\arccos(1-u)\]


```{r, PROBLEM 1}

set.seed(123)
u<-runif(n=10000)
x<-acos(1-u)
y<-seq(from=0,to=pi/2,0.01)
f<-sin(y)
hist(x,freq = F)
lines(y,f,col="red")
```

##### PART B: REJECTION SAMPLING METHOD
```{r}
M<-2 #scaling
n<-10000 #number of iterations
final_sample<-rep(NA,n)
accept<-0
while (accept<n){
    y<-runif(1,0,pi/2) #get one random number between 0 and pi/2 as the value to input in f(x)
    u<-runif(1,0,1)  #another random number between 0 and 1 to compare
    if (M*u<=sin(y)){  #if the value of M*u is less or equal than the target distribution, then
        
        accept<-accept+1 #increase the counter by 1 to be able to continue storing in final_sample
        final_sample[accept]<-y #store the random number that was evaluated in f(x)  
    }
}
p<-seq(from=0,to=pi/2,0.01)
w<-sin(p)

hist(final_sample,freq=F)
lines(p,w,col="red")

```


***

#### PROBLEM 2
##### PART A

$p(x)=\lambda x^{-\lambda -1}$ therefore:
\[\int_1^{x} w^{-\lambda -1}~dw\] which yields:
\[F(x)=1-x^{-\lambda}\]
Using the inversion method:
\[F^{-1}=\frac{1}{\sqrt[\lambda]{1-u}}\]

```{r, PROBLEM 2}
lambda<-4
lambda_fun<-function(lambda,n=1){
u<-runif(n)
x<-1/((1-u)^(1/lambda))
return(x)
}
y=seq(1,6,0.2)
fy=(lambda*y^(-lambda-1))
hist(lambda_fun(lambda,100),freq=F)
lines(y,fy,col="red")
```

##### PART B

According to the definition of expectation:
$$E(X)=\int_{-\infty}^{+\infty}xP(x)dx $$
The estimation $E(X^{2.5})$, where $\lambda>2.5$ corresponds to:
$$E(X^{2.5})=\lambda\int_{1}^{\infty}x^{2.5}x^{-\lambda-1}dx=\left[\frac{\lambda x^{1.5-\lambda+1}}{1.5-\lambda+1}\right]^{\infty}_{1}=0-\frac{\lambda}{2.5-\lambda}=\frac{3}{3-2.5}=6 $$

Using R to verify this result:
```{r PROBLEM 2b}
lambda<-3
delta<-2.5
f<-function(x){x^delta*lambda*x^(-lambda-1)}
val<-integrate(f,1,Inf)
return(val)



```
***

#### PROBLEM 3

Function is \[f(x)\propto\exp (-x^{4}/12)\]

##### STEP 1

under the transformation $x=y^{1/n}$, where $dx=\frac{1}{n}y^{\frac{1}{n}-1} dy$, and substituting accordingly, the transformed density becomes:
\[f(y)=\frac{1}{n}\int_0^{\infty}{\rm e}^{-ky}{y}^{\frac{1}{n}-1}dy\] 
Where the constant $\frac{1}{n}$ can be omitted because of proportionality.

Comparing with the density of Gamma, $\alpha=\frac{1}{n}$ and $\lambda=k$

```{r}
alpha=1/4
lambda=1/12
n<-1E5
y<-rep(NA,n)
y<-rgamma(n,shape=alpha, rate=lambda)
z<-y^(alpha)
#the following lines are taken from the in-class exercise slides
hist(z, breaks = 30, freq = F, col = rgb(0.75,0.4,0.1,0.5)) # z is your sample
target <- function(x){exp(-lambda*x^(1/alpha))/
integrate(function(x) exp(-lambda*x^(1/alpha)),0,Inf)$value}
curve(target,lwd=2,add=T)
````
