---
title: STAT 5443, HW3
author: Ariel Mundo
date: "`r Sys.Date()`"
output:
    prettydoc::html_pretty:
    theme: cayman
    #css: styles.css
---
<style type="text/css">

p {
   font-size: 14px;
}
code.r{
  font-size: 14px;
}
pre {
  font-size: 14px
}
</style>



```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(prettydoc)
```

## PROBLEM 1
### GIBBS SAMPLER

This section contains parts 1-4 from PROBLEM 1.
```{r, GIBBS SAMPLER}
library(plyr)
rm(list = ls())
set.seed(123)
alpha=2
beta_o=6.4
n=74
s_1=16
niter=1000
theta_init=s_1/n
s_init=s_1/n
  theta=rep(NA,niter)
  s=rep(NA,niter)
  theta[1] = theta_init
  s[1]=s_init
  for(i in 2:niter){
    s[i] = rbinom(1,size=n,prob=theta[i-1])
    theta[i] = rbeta(1,alpha+s[i],beta_o+n-s[i])
  }
  pmf<-count(s)
  pmf$vals<-pmf$freq/niter
  
plot(theta,type="l", main='Traceplot for theta',xlab='iterations')
plot(s,type="l", main='Traceplot for s',xlab='iterations')
plot(x=pmf$x,y=pmf$vals, main='approximate pmf')
hist(s,freq=FALSE, main='Histogram for s')
print(paste('posterior median of theta:',median(theta), 'MLE estimate',s_1/n))
```
<div class = "blue">
 *The estimated posterior mean of* $\theta$ *is indeed to the maximum likelihood estimate of s/n .*
</div>
http://www.stat.cmu.edu/~brian/463-663/week10/Chapter%2004.pdf
https://bookdown.org/rdpeng/advstatcomp/gibbs-sampler.html

### PART 5 OF PROBLEM 1

This section measures the sensitivity of the posterior median to the choice of initial values. For this part, the previous Gibbs sampler will be transformed in a function, and the initial ratio $s/n$ will be randomized and the median $s$ and $\theta$ will be plotted for all iterations.

```{r, PROBLEM 1 PART 5}

binomial_beta_gibbs<-function(theta_init,s_init,iterations){
alpha=2
beta_o=6.4
n=74
  theta_init=s_1/n
s_init=s_1/n
  theta=rep(NA,iterations)
  s=rep(NA,iterations)
  theta[1] = theta_init
  s[1]=s_init
  for(i in 2:iterations){
    s[i] = rbinom(1,size=n,prob=theta[i-1])
    theta[i] = rbeta(1,alpha+s[i],beta_o+n-s[i])
  }
  pmf<-count(s)
  pmf$vals<-pmf$freq/niter
 return(cbind(median(s),median(theta)))
}
#v<-seq(0,10,0.1)
v<-runif(10)
test<-matrix(NA,2,length(v))
for (j in 1:length(v)){
test[,j]<-binomial_beta_gibbs(v[j],v[j],1000)
}
plot(test[2,], type='l', main='median s from variable initial values') #plotting the median values of all iterations
plot(test[1,], type='l', main='median theta from variable initial values') #plotting the median values of all iterations

```

***

### PROBLEM 2


