---
title: STAT 5443, Midterm
author: Ariel Mundo
date: 10/12/2020
output:
    prettydoc::html_pretty:
    theme: cayman
    #css: styles.css
---
<style type="text/css">

p {
   font-size: 12px;
}
code.r{
  font-size: 12px;
}
pre {
  font-size: 12px
}
</style>



```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(prettydoc)
```
 ### PROBLEM 1
 
 #### Part A
For this problem, rejection sampling can be used.
```{r, PROBLEM 1}
M<-6 #scaling
n<-1E5 #number of iterations
final_sample<-rep(NA,n)
accept<-0
while (accept<n){
    y<-runif(1,0,1) #get one random number between 0 and pi/2 as the value to input in f(x)
    u<-runif(1,0,1)  #another random number between 0 and 1 to compare
    if (M*u<=(1+sqrt(y)*exp(-(y/2+sqrt(y))))/(2*sqrt(y))){  #if the value of M*u is less or equal than the target distribution, then
        
        accept<-accept+1 #increase the counter by 1 to be able to continue storing in final_sample
        final_sample[accept]<-y #store the random number that was evaluated in f(x)  
    }
}
p<-seq(from=0,to=1,0.01)
w<-(1+sqrt(p)*exp(-(p/2+sqrt(p))))/(2*sqrt(p))

hist(final_sample,freq=F)
lines(p,w,col="red")

```
#### Part B
 
 Using Monte Carlo integration:
 
```{r}
set.seed(1245)
n=1E4
f<-function(x){
        log(x)*((1+sqrt(x))/(2*sqrt(x)))*exp(-(x/2+sqrt(x))) #function
}
val=f(runif(n))
Expectation<-mean(val) #calculate the expectation
Std_err=sqrt(var(val)/n) #calculate standard error
CI=1.96*Std_err #confidence interval corresponding to 95%
print(paste('Expectation:',Expectation))
print(paste('95% Confidence interval:[',Expectation-CI,",",Expectation+CI,']'))
```
    Edelman's answer was -1.68788 and while the simple Monte Carlo does not match the value exactly, it provides a confidence interval that does contain the true value.


#### PROBLEM 3
The algorithm will correspond to the following in R:
```{r}
test<-function(n){
  if (n>1){
    print('still going')
   test(n/2)
    test(n/2)
  }
 
}
```
The first comparison (if) will print the text once. Next, the function is recursively called twice by dividing n/2, and will print the text each time it is called until succesive divisions of  n/2 result in a value <1. This is done two times in each ocassion as the recursion happens twice. Therefore:
$T(n)=2T(\frac{n}{2})+1$, where $a=2,b=2,d=0$
According to the Master Theorem, as in this case $d<\log_b(a)$, this results in:
$$O(n^{\log_22})=n $$
Therefore, the algorithm will print $\Theta(n)= n$ lines.
