---
title: "Practice problems"
author: "Ariel Mundo"
date: "9/30/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Problem 1

This would be the case of Monte Carlo integration by the sample-mean method.
This would involve to generate a large number of samples from the uniform distribution in the interval $ (-\pi,\pi)$

```{r PROBLEM 1}
a=-pi
b=pi
n=1E5
x=runif(n,min=a,max=b)
f<-(exp(x)-exp(-x))/2
int<-(b-a)*mean(f)


```


## Problem 3

Devise a recursive algorithm to find the sum of the first n odd positive integers.
Using the odd number theorem, which states that:

$$\sum_{i=1}^{n}(2i-1)=n^{2}$$
Pseudocode:

read number $n$

calculate $n*n$

print result

There is a single call to the algorithm even if the number is large, therefore the time is constant and $O(1)$
```{r}
n=10
j=rep(0,n)
j[1]<-1
for (i in 1:n-1){ #create array of odd numbers
j[i+1]<-2*i+1
}
total<-sum(j)
print(total)
```
## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

